---
title       : Demo Analysis with WB Procurement Data
subtitle    : 
author      : Tommy Shen
job         : Data Scientist
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
---

## Demos of Data Analytics & Metrics

Sample visualizations with World Bank Procurement data from 1995 to 2014.


```{r echo=FALSE, cache=TRUE}
wbdat <- read.csv("wb_data.csv")

## Get useable names
names(wbdat) <- gsub("\\.", "_", tolower(names(wbdat)))
names(wbdat) <- gsub("__", "_", names(wbdat))


wbdat$amount <- as.integer(gsub(",", "", wbdat$amount))
wbdat$count <- 1

```

---

## Top 10 suppliers by contracts

```{r, echo=FALSE, results='asis'}
suppressMessages(library(pander))
topsupplier_amt <- aggregate(cbind(amount,count) ~ supplier, data=wbdat, sum)

a <- head(topsupplier_amt[order(-topsupplier_amt$count),], n=10)
print("By Contracts")
pander(a[,c(1,3)], style = 'rmarkdown', split.cells = Inf)

# a <- head(topsupplier_amt[order(-topsupplier_amt$amount),], n=10)
#pander(a[,1:2], style = 'rmarkdown', split.cells = Inf)

```

---

## Sector distribution of contract awards
```{r, results='asis'}
pander(aggregate(count ~ major_sector, data=wbdat, sum), style = 'rmarkdown', split.cells = Inf)
```

---

## Distribution of contract awards by procurement category

```{r, results='asis'}
pander(aggregate(count ~ proc_categ, data=wbdat, sum), style = 'rmarkdown', split.cells = Inf)
```

---
## Award Ratios by Procurement Method
```{r fig.height=7, fig.width=12, echo=FALSE}
library(RColorBrewer)
proc_meth <- aggregate(cbind(amount,count) ~ proc_meth, data=wbdat, sum)

proc_meth$amount <- proc_meth$amount/sum(proc_meth$amount)
proc_meth$count <- proc_meth$count/sum(proc_meth$count)

par(mar=c(5,5,2,2))
barplot(as.matrix(proc_meth[,2:3]),ylim=c(0, 2 + 3), col=colorRampPalette(brewer.pal(8, "Accent"))(16), horiz=TRUE, main="Award Ratios by Procurement Method", legend.text=proc_meth$proc_meth, args.legend = list(y = 2 + 3, bty = "n", ncol=3))

```

---

## Award Ratios by Supplier Country

```{r fig.height=7, fig.width=12, echo=FALSE}
proc_ctry <- aggregate(cbind(amount,count) ~ supp_ctry, data=wbdat, sum)

proc_ctry$amount <- proc_ctry$amount/sum(proc_ctry$amount)
proc_ctry$count <- proc_ctry$count/sum(proc_ctry$count)


par(mar=c(5,5,2,2))
barplot(as.matrix(proc_ctry[,2:3]),ylim=c(0, 2 + 3), col=colorRampPalette(brewer.pal(8, "Accent"))(16), horiz=TRUE, main="Award Ratios by Supplier Country", legend.text=proc_ctry$supp_ctry, args.legend = list(y = 2 + 3, bty = "n", ncol=6))

```

---

## Distribution of Spending

```{r results='asis', echo=FALSE}
library(plyr)
library(rCharts)

a <- ddply(wbdat, .(proc_categ, major_sector), summarize, amount = sum(amount, na.rm=TRUE))
sec_plot <- nPlot(amount ~ major_sector, group = 'proc_categ', data=a, type='multiBarChart')
print(sec_plot)

```


---

## Time Series by Procurement Category

```{r results='asis', echo=FALSE}
categ_series <- aggregate(amount ~ fiscal_year + proc_categ, data=wbdat, sum)
h1 <- Highcharts$new()
h1$chart(type = "spline")
for(i in unique(categ_series$proc_categ)){
  a <- subset(categ_series, proc_categ==i)
  b <- sapply(sort(unique(categ_series$fiscal_year)), function(y){
    if(y %in% a$fiscal_year) return(a$amount[a$fiscal_year==y])
    else return(NA)
  })
  h1$series(data = b, name=i, dashStyle = "shortdash")
}
h1$xAxis(categories = sort(unique(categ_series$fiscal_year)))
h1$legend(symbolWidth = 80)
h1
```


## Demos of Data Analytics & Metrics

Sample visualizations with World Bank Procurement data from 1995 to 2014.


```{r echo=FALSE, cache=TRUE}
library(plyr)
library(rCharts)
wbdat <- read.csv("./data/wb_data.csv")

## Get useable names
names(wbdat) <- gsub("\\.", "_", tolower(names(wbdat)))
names(wbdat) <- gsub("__", "_", names(wbdat))

wbdat$amount <- as.integer(gsub(",", "", wbdat$amount))
wbdat$count <- 1

```

---

## Top 10 suppliers by contracts

```{r, echo=FALSE, results='asis'}
suppressMessages(library(pander))
topsupplier_amt <- aggregate(cbind(amount,count) ~ supplier, data=wbdat, sum)

a <- head(topsupplier_amt[order(-topsupplier_amt$count),], n=10)
print("By Contracts")
pander(a[,c(1,3)], style = 'rmarkdown', split.cells = Inf)

# a <- head(topsupplier_amt[order(-topsupplier_amt$amount),], n=10)
#pander(a[,1:2], style = 'rmarkdown', split.cells = Inf)

```

---

## Sector distribution of contract awards
```{r, results='asis'}
pander(aggregate(count ~ major_sector, data=wbdat, sum), style = 'rmarkdown', split.cells = Inf)
```

---

## Distribution of contract awards by procurement category

```{r, results='asis'}
pander(aggregate(count ~ proc_categ, data=wbdat, sum), style = 'rmarkdown', split.cells = Inf)
```

---
## Award Ratios by Procurement Method
```{r fig.height=7, fig.width=12, echo=FALSE}
library(RColorBrewer)
proc_meth <- aggregate(cbind(amount,count) ~ proc_meth, data=wbdat, sum)

proc_meth$amount <- proc_meth$amount/sum(proc_meth$amount)
proc_meth$count <- proc_meth$count/sum(proc_meth$count)

par(mar=c(5,5,2,2))
barplot(as.matrix(proc_meth[,2:3]),ylim=c(0, 2 + 3), col=colorRampPalette(brewer.pal(8, "Accent"))(16), horiz=TRUE, main="Award Ratios by Procurement Method", legend.text=proc_meth$proc_meth, args.legend = list(y = 2 + 3, bty = "n", ncol=3))

```

---

## Award Ratios by Supplier Country

```{r fig.height=7, fig.width=12, echo=FALSE}
proc_ctry <- aggregate(cbind(amount,count) ~ supp_ctry, data=wbdat, sum)

proc_ctry$amount <- proc_ctry$amount/sum(proc_ctry$amount)
proc_ctry$count <- proc_ctry$count/sum(proc_ctry$count)


par(mar=c(5,5,2,2))
barplot(as.matrix(proc_ctry[,2:3]),ylim=c(0, 2 + 3), col=colorRampPalette(brewer.pal(8, "Accent"))(16), horiz=TRUE, main="Award Ratios by Supplier Country", legend.text=proc_ctry$supp_ctry, args.legend = list(y = 2 + 3, bty = "n", ncol=6))

```

---

## Detailed look at country/sector

```{r fig.height=7, fig.width=12, echo=FALSE, results='asis'}
library(plyr)
library(rCharts)
countries <- aggregate(amount ~ supp_ctry, data=wbdat, sum)
countries <- countries[order(-countries$amount),]
countries <- head(countries, n=5)

proc_ctry <- aggregate(amount ~ supp_ctry + major_sector, data=wbdat, sum)
proc_ctry <- subset(proc_ctry, supp_ctry %in% countries$supp_ctry)

ctry_plot <- nPlot(amount ~ supp_ctry, group = 'major_sector', data=proc_ctry, type='multiBarChart')
ctry_plot

```
---

## Distribution of Spending

```{r results='asis', echo=FALSE}
a <- ddply(wbdat, .(proc_categ, major_sector), summarize, amount = sum(amount, na.rm=TRUE))
sec_plot <- nPlot(amount ~ major_sector, group = 'proc_categ', data=a, type='multiBarChart')
sec_plot

```


---


## Time Series (Top 5 Sectors)

```{r results='asis', echo=FALSE}

sectors <- aggregate(amount ~ major_sector, data=wbdat, sum)
sectors <- sectors[order(-sectors$amount),]
sectors <- head(sectors, n=5)


categ_series <- aggregate(amount ~ fiscal_year + major_sector, data=wbdat, sum)
categ_series <- subset(categ_series, major_sector %in% sectors$major_sector)
h1 <- Highcharts$new()
h1$chart(type = "spline")
for(i in unique(categ_series$major_sector)){
  a <- subset(categ_series, major_sector==i)
  b <- sapply(sort(unique(categ_series$fiscal_year)), function(y){
    if(y %in% a$fiscal_year) return(a$amount[a$fiscal_year==y])
    else return(NA)
  })
  h1$series(data = b, name=i, dashStyle = "shortdash")
}
h1$xAxis(categories = sort(unique(categ_series$fiscal_year)))
h1$legend(symbolWidth = 80)
h1
```

---

## Time Series by Procurement Category

```{r results='asis', echo=FALSE}
categ_series <- aggregate(amount ~ fiscal_year + proc_categ, data=wbdat, sum)
h1 <- Highcharts$new()
h1$chart(type = "spline")
for(i in unique(categ_series$proc_categ)){
  a <- subset(categ_series, proc_categ==i)
  b <- sapply(sort(unique(categ_series$fiscal_year)), function(y){
    if(y %in% a$fiscal_year) return(a$amount[a$fiscal_year==y])
    else return(NA)
  })
  h1$series(data = b, name=i, dashStyle = "shortdash")
}
h1$xAxis(categories = sort(unique(categ_series$fiscal_year)))
h1$legend(symbolWidth = 80)
h1
```
